### 2022-01-28 chapters/thesis_outline.tex
update

The work described in the present thesis intended to XXXXX.

The thesis comprises X chapters, organised as follows:

\textbf{Chapter 1} corresponds to the general introduction that highlights the ...

\textbf{Chapter 2} is dedicated to ...

\textbf{Chapter 3} consists in ...

\textbf{Chapter 4}

\textbf{Chapter 5}

\textbf{Chapter 6}

\textbf{Chapter 7}

\textbf{Chapter 8} corresponds to the general discussion. In this chapter is provided a
summary of the main results obtained in this thesis and its integrated discussion.

\textbf{Chapter 9} Contains the main conclusions driven from this work. It also includes perspectives for future work.