### 2022-02-20 main.tex

%-----------------------------------------------------------------
% Preamble
%------------------------------------------------------------------
\documentclass[12pt,a4paper,twoside,openright]{book}
\usepackage{mathptmx} % times no roman-like font
\usepackage[left=30mm, right=25mm, top=25mm, bottom=25mm]{geometry} % margins
\pagestyle{fancy}
\renewcommand{\baselinestretch}{1.15}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{figures/}}
\def\blankpage{%
      \clearpage%
      \thispagestyle{empty}%
      \null%
      \clearpage}

%-----------------------------------------------------------------
% Packages
%------------------------------------------------------------------
\usepackage{setspace}
\usepackage{fancyhdr}
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\thechapter. #1 }}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\usepackage{titlesec}
\fancyhf{}
\fancyhead[RO]{\bfseries\rightmark}
\fancyhead[LE]{\bfseries\leftmark}
\setlength{\headheight}{15pt}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\headheight}{0.5pt}
\fancypagestyle{plain}{
  \fancyhead{}
  \renewcommand{\headrulewidth}{0pt}
}
\usepackage{hyperref} % To add urls
\usepackage{graphicx} % To add images
\usepackage[nottoc]{tocbibind}
\usepackage[backend=biber,style=authoryear,maxcitenames=1,autocite=inline,natbib=true,sorting=nty]{biblatex} % References
\addbibresource{references.bib}
\usepackage{indentfirst} % Add indentation to first paragraph
\setlength{\parskip}{1em}
\setcounter{tocdepth}{4}
\setcounter{secnumdepth}{4}
\usepackage{enumitem}
\usepackage[font=footnotesize]{caption}
\usepackage[toc,page]{appendix}
\usepackage{pdfpages}
\usepackage{fmtcount}
\usepackage{url}
\usepackage{amsmath}
\usepackage{bm}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage{float}
\raggedbottom
\usepackage{subcaption}
\usepackage{eurosym}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[acronym]
\makeatletter
\def\cleardoublepage{\clearpage\if@twoside \ifodd\c@page\else
    \hbox{}
    \thispagestyle{plain}
    \newpage
    \if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother \clearpage{\pagestyle{plain}\cleardoublepage}

\providecommand{\keywords}[1]{\textbf{Keywords:} #1}
\providecommand{\keywordsP}[1]{\textbf{Palavras chave:} #1}
\newcommand{\secref}[1]{\autoref{#1}. \nameref{#1}}

%-----------------------------------------------------------------
% Begin Document
%------------------------------------------------------------------

\begin{document}
\begin{sloppy}
\frontmatter

%-----------------------------------------------------------------
% Cover
%------------------------------------------------------------------
\input{titlepage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Folha de rosto
%------------------------------------------------------------------
\input{coverpage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Disclaimer
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{\fill}
As opiniões expressas nesta publicação são da exclusiva responsabilidade do seu autor.
\blankpage

%-----------------------------------------------------------------
% Dedication
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{3cm}
\begin{center}
\emph{"The greatest adventure is what lies ahead.\\
Today and tomorrow are yet to be said.\\
The chances, the changes are all yours to make.\\
The mould of your life is in your hands to break."} \\
\end{center} 
\begin{flushright}-J. R. R. Tolkien, The Hobbit\end{flushright}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Acknowledgements
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Acknowledgements}
\chapter*{Acknowledgements}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Summary
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Summary}
\chapter*{Summary}

\keywords{one, two, three, four, five}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Resumo
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Resumo}
\chapter*{Resumo}

\keywords{um, dois, três, quatro, cinco}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Thesis Outline
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Thesis Outline}
\chapter*{Thesis Outline}
\input{chapters/thesis_outline}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Abbreviations
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Abbreviation}
\chapter*{Abbreviation}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Table of Contents
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Table of Contents}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoftables
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoffigures
\clearpage \thispagestyle{empty}\mbox{}\clearpage

\mainmatter
%\pagenumbering{arabic} %numeros em numeração árabe

%-----------------------------------------------------------------
% Introduction
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{General Introduction\label{ch:introduction}}
\thispagestyle{empty}
\newpage
\begin{refsection}
\include{chapters/introduction}
\newpage
\printbibliography
\end{refsection}

%-----------------------------------------------------------------
% Conclusion
%-----------------------------------------------------------------
\newpage
\chapter{Conclusion \label{ch:conclusion}}
\thispagestyle{empty}
\newpage
\include{chapters/conclusion}

%-----------------------------------------------------------------
% Appendix
%-----------------------------------------------------------------
\appendix
\chapter{Appendix}
\include{chapters/appendix}
%\includepdf[pages={1-}]{resources/papers/s41598-018-31873-w.pdf}
%\includepdf[pages={1-}]{resources/papers/dkz363.pdf}
%\includepdf[pages={1-}]{resources/papers/mgen000328.pdf}

\end{sloppy}
\end{document}