### 2022-02-28 chapters/accronims.tex

Global Burden of Disease (GBD)
World Health Organization (WHO)
acute  respiratory  syndrome  coronavirus  2(SARS-CoV-2)
Middle East respiratory syndrome (MERS)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)
PubMed Central\textsuperscript{\small\textregistered} (PMC)
Public health surveillance systems (PHSS)
dideoxynucleotide triphosphate (ddNTP) 
deoxynucleotide triphosphate (dNTP)
single molecule real-time sequencing (SMRT)
Whole genome sequencing (WGS)
ribosomal RNA (rRNA)
Operational Taxonomic Units (OTUs)
Shiga toxin-producing \textit{Escherichia coli} (STEC)
National Institute of Standards and Technology (NIST)

R-PCR real time