### 2022-03-26 chapters/acconyms.tex
start list of abbreviations

\begin{acronym}[MPC] % Give the longest label here so that the list is nicely aligned
\acro{GBD}{Global Burden of Disease}
\acro{WHO}{World Health Organization}
\acro{SARS-CoV-2}{Acute  Respiratory  Syndrome  Coronavirus  2}
\acro{MERS}{Middle East Respiratory Syndrome}
\acro{GPP}{Global Priority Pathogens}
\acro{AMR}{Antimicrobial Resistance}

\end{acronym}

T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)
PubMed Central\textsuperscript{\small\textregistered} (PMC)
Public health surveillance systems (PHSS)
dideoxynucleotide triphosphate (ddNTP) 
deoxynucleotide triphosphate (dNTP)
single molecule real-time sequencing (SMRT)
Whole genome sequencing (WGS)
ribosomal RNA (rRNA)
Operational Taxonomic Units (OTUs)
Shiga toxin-producing \textit{Escherichia coli} (STEC)
National Institute of Standards and Technology (NIST)
reverse transcription-PCR (RT-PCR)
core-genome/whole genome MLST (cg/wg MLST)
overlap-layout consensus (OLC)
Cerebrospinal fluid (CSF)

chap 2 
Shotgun metagenomics (SMg) *** Check in introduction
whole-genome sequence (WGS)
Federation of Dutch Medical Scientific Societies (FDMSS)
Univerisity Medical Center Groningen (UMCG)

R-PCR real time