### 2022-03-26 chapters/acconyms.tex
continue

\begin{acronym}[MPC] % Give the longest label here so that the list is nicely aligned
\acro{bp}{basepairs}
\acro{dNTP}{deoxynucleotide triphosphate}
\acro{ddNTP}{dideoxynucleotide triphosphate}
\acro{qPCR}{Real-Time Quantitative PCR}
\acro{rRNA}{ribossomal RNA}

\acro{AMR}{Antimicrobial Resistance}
\acro{GBD}{Global Burden of Disease}
\acro{GPP}{Global Priority Pathogens}
\acro{LFIA}{Lateral Flow Immunoassays}
\acro{MERS}{Middle East Respiratory Syndrome}
\acro{MLST}{Multilocus Sequence Typing}
\acro{PCR}{Polymerase Chain Reaction}
\acro{PFGE}{Pulse Field Gel Electrophoresis}
\acro{PHSS}{Public Health Surveillance Systems}
\acro{PMC}{PubMed Central\textsuperscript{\small\textregistered}}
\acro{SARS-CoV-2}{Acute  Respiratory  Syndrome  Coronavirus  2}
\acro{WHO}{World Health Organization}

\end{acronym}

single molecule real-time sequencing (SMRT)
Whole genome sequencing (WGS)
ribosomal RNA (rRNA)
Operational Taxonomic Units (OTUs)
Shiga toxin-producing \textit{Escherichia coli} (STEC)
National Institute of Standards and Technology (NIST)
reverse transcription-PCR (RT-PCR)
core-genome/whole genome MLST (cg/wg MLST)
overlap-layout consensus (OLC)
Cerebrospinal fluid (CSF)
core-genome/whole genome MLST (cg/wg MLST)

chap 2 
Shotgun metagenomics (SMg) *** Check in introduction
whole-genome sequence (WGS)
Federation of Dutch Medical Scientific Societies (FDMSS)
Univerisity Medical Center Groningen (UMCG)

R-PCR real time