### 2022-04-29 chapters/chapter_8.tex
update

\mbox{}\\
\vspace{8cm}

This chapter is a reproduction of the following publication:

B. C. L. van der Putten, C. I. Mendes, B. M. Talbot, J. de Korne-Elenbaas, R. Mamede, P. Vila-Cerqueira, L. P. Coelho, C. A. Gulvik, L. S. Katz and The ASM NGS 2020 Hackathon participants. Software testing in microbial bioinformatics: a call to action. Microbial Genomics, Volume 8, Issue 3, March 2020. DOI: \url{https://doi.org/ 10.1099/mgen.0.000790}