### 2022-05-16 chapters/chapter_6.tex
Revert "start discussion"
This reverts commit edfc8947b90b00c7968d6707c7048c0254029344.

\mbox{}\\
\vspace{8cm}

This chapter is a reproduction of the following submitted manuscript for publication in eLife:

hAMRonization