### 2022-05-16 chapters/chapter_6.tex
update with working version

\mbox{}\\
\vspace{8cm}

This chapter is a reproduction of the following submitted manuscript for publication in eLife:

C. I. Mendes, F. Maguire, A. Manuele, D. Fornika, S. Tausch, T. Le-Viet, J. Phelan, C. Meehan, A. Raphenya, B. Alcock, E. Culp, A. McArthur, M. Feldgarden, G. Tyson, M. Galas, J. Campos, A. Witney, D. Aanensen, A. Black, L. Katz, P. Oluniyi, I. Olawoye, R. Timme, A. T. Vasconcelos, A. Page, D. MacCannell, E. Griffiths. 
hAMRonization: Enhancing antimicrobial resistance prediction using PHA4GE standards and specification

My contribution to this publication included the development of the antimicrobial detection con-
textual data specification package, including it’s conversion and availability in a machine-
applicable JSON format. It also includes the design and implementation of the hAMRonization package, results analysis, manuscript production and editing.