### 2022-05-23 tables/chapter_4/suptable2.tex

\begin{longtable}{@{}lllll@{}}
\caption{Run accession ID, BioProject SRA Study ID, source and organism present for each sample of the negative control dataset (ZKV – zika virus, CHIKV – chikungunya virus, YFV – yellow fever virus).}
\label{tab:chap4_s2}\\
\toprule
Run ID     & Bioproject  & SRA Study & Source                & Organism \\ \midrule
SRR8031152 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8062732 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8031153 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8063606 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8063603 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8063605 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8031155 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8031154 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8063604 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR8062733 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985391 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985394 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985620 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & CHIKV    \\
SRR7985390 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985392 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985621 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & CHIKV    \\
SRR5179639 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179637 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179646 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR7985389 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR7985622 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & CHIKV    \\
SRR7985619 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & CHIKV    \\
SRR5179667 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179653 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR7985393 & PRJNA494391 & SRP163225 & Shotgun Metagenomic   & ZKV      \\
SRR5179638 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179636 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179666 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179650 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179649 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179643 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179635 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179645 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179642 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179644 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179647 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179641 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179640 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179652 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179648 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR5179651 & PRJNA361543 & SRP096859 & Amplicon Metagenomics & YFV      \\
SRR9020503 & PRJNA541092 & SRP195668 & Amplicon Metagenomics & CHIKV    \\
SRR9020505 & PRJNA541093 & SRP195669 & Amplicon Metagenomics & CHIKV    \\
SRR9020506 & PRJNA541094 & SRP195670 & Amplicon Metagenomics & CHIKV    \\
SRR9020509 & PRJNA541095 & SRP195671 & Amplicon Metagenomics & CHIKV    \\
SRR9020511 & PRJNA541096 & SRP195672 & Amplicon Metagenomics & CHIKV    \\
SRR9020513 & PRJNA541097 & SRP195673 & Amplicon Metagenomics & CHIKV    \\
SRR9020514 & PRJNA541098 & SRP195674 & Amplicon Metagenomics & CHIKV    \\
SRR9020516 & PRJNA541099 & SRP195675 & Amplicon Metagenomics & CHIKV    \\
SRR9020518 & PRJNA541100 & SRP195676 & Amplicon Metagenomics & CHIKV    \\
SRR9020520 & PRJNA541101 & SRP195677 & Amplicon Metagenomics & CHIKV    \\
SRR9020521 & PRJNA541102 & SRP195678 & Amplicon Metagenomics & CHIKV    \\
SRR9020523 & PRJNA541103 & SRP195679 & Amplicon Metagenomics & CHIKV    \\
SRR9020525 & PRJNA541104 & SRP195680 & Amplicon Metagenomics & CHIKV    \\
SRR9020527 & PRJNA541105 & SRP195681 & Amplicon Metagenomics & CHIKV    \\
SRR9020529 & PRJNA541106 & SRP195682 & Amplicon Metagenomics & CHIKV    \\
SRR9020530 & PRJNA541107 & SRP195683 & Amplicon Metagenomics & CHIKV    \\
SRR9020532 & PRJNA541108 & SRP195684 & Amplicon Metagenomics & CHIKV    \\
SRR9020534 & PRJNA541109 & SRP195685 & Amplicon Metagenomics & CHIKV    \\
SRR9020537 & PRJNA541110 & SRP195686 & Amplicon Metagenomics & CHIKV    \\
SRR9020539 & PRJNA541111 & SRP195687 & Amplicon Metagenomics & CHIKV    \\
SRR9020541 & PRJNA541112 & SRP195688 & Amplicon Metagenomics & CHIKV    \\
SRR9020542 & PRJNA541113 & SRP195689 & Amplicon Metagenomics & CHIKV    \\
SRR9020504 & PRJNA541114 & SRP195690 & Amplicon Metagenomics & CHIKV    \\
SRR9020507 & PRJNA541115 & SRP195691 & Amplicon Metagenomics & CHIKV    \\
SRR9020508 & PRJNA541116 & SRP195692 & Amplicon Metagenomics & CHIKV    \\
SRR9020510 & PRJNA541117 & SRP195693 & Amplicon Metagenomics & CHIKV    \\
SRR9020512 & PRJNA541118 & SRP195694 & Amplicon Metagenomics & CHIKV    \\
SRR9020515 & PRJNA541119 & SRP195695 & Amplicon Metagenomics & CHIKV    \\
SRR9020517 & PRJNA541120 & SRP195696 & Amplicon Metagenomics & CHIKV    \\
SRR9020519 & PRJNA541121 & SRP195697 & Amplicon Metagenomics & CHIKV    \\
SRR9020522 & PRJNA541122 & SRP195698 & Amplicon Metagenomics & CHIKV    \\
SRR9020524 & PRJNA541123 & SRP195699 & Amplicon Metagenomics & CHIKV    \\
SRR9020526 & PRJNA541124 & SRP195700 & Amplicon Metagenomics & CHIKV    \\
SRR9020528 & PRJNA541125 & SRP195701 & Amplicon Metagenomics & CHIKV    \\
SRR9020531 & PRJNA541126 & SRP195702 & Amplicon Metagenomics & CHIKV    \\
SRR9020533 & PRJNA541127 & SRP195703 & Amplicon Metagenomics & CHIKV    \\
SRR9020535 & PRJNA541128 & SRP195704 & Amplicon Metagenomics & CHIKV    \\
SRR9020536 & PRJNA541129 & SRP195705 & Amplicon Metagenomics & CHIKV    \\
SRR9020538 & PRJNA541130 & SRP195706 & Amplicon Metagenomics & CHIKV    \\
SRR9020540 & PRJNA541131 & SRP195707 & Amplicon Metagenomics & CHIKV    \\
SRR7369225 & PRJNA47661  & SRP150883 & Shotgun Metagenomic   & ZKV      \\
SRR7369226 & PRJNA47661  & SRP150883 & Shotgun Metagenomic   & ZKV      \\
SRR6505781 & PRJNA431343 & SRP131290 & Shotgun Metagenomic   & ZKV      \\
SRR8260975 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8260976 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8260977 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8260978 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8260979 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8260980 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261322 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261325 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261326 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261329 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261330 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261331 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261332 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261333 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261335 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261336 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261338 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261341 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261342 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261343 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261345 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261346 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261347 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261348 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261352 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261353 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261354 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261355 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261356 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261359 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261360 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261361 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261362 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261364 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261365 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261366 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261367 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261369 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261402 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261404 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261407 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261411 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261412 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261413 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261415 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261416 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV      \\
SRR8261417 & PRJNA314889 & SRP162155 & Shotgun Metagenomic   & ZKV  \bottomrule   
\end{longtable}