### 2022-05-23 tables/chapter_7/table1.tex

\begin{table}[h!]
\caption{Ontologies implemented in the PHA4GE SARS-CoV-2 specification.}
\label{tab:ch7_table1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}ll@{}}
\toprule
Ontology$^1$                                    & Link                                           \\ \midrule
BRENDA Tissue Ontology (BTO)                 & https://obofoundry.org/ontology/bto.html       \\
Cell Line Ontology (CLO)                     & https://obofoundry.org/ontology/clo.html       \\
Environmental conditions,   treatments and exposures ontology (ECTO) & https://obofoundry.org/ontology/ecto.html   \\
Environment Ontology (ENVO)                  & https://obofoundry.org/ontology/envo.html      \\
Food Ontology (FoodOn)                       & https://obofoundry.org/ontology/foodon.html    \\
Gazetteer Ontology (GAZ)                     & https://obofoundry.org/ontology/gaz.html       \\
Gender, Sex, and Sexual Orientation Ontology (GSSO)                  & https://obofoundry.org/ontology/gsso.html   \\
Genomic Epidemiology Ontology (GenEpiO)      & https://obofoundry.org/ontology/genepio.html   \\
Genomics Cohorts Knowledge Ontology (GECKO)  & https://obofoundry.org/ontology/gecko.html     \\
Human Disease Ontology (DOID)                & https://obofoundry.org/ontology/doid.html      \\
Human Phenotype Ontology (HP)                & https://obofoundry.org/ontology/hp.html        \\
Mammalian Phenotype Ontology (MP)            & https://obofoundry.org/ontology/mp.html        \\
Measurement Method Ontology (MMO)            & https://obofoundry.org/ontology/mmo.html       \\
Mondo Disease Ontology (MONDO)               & https://obofoundry.org/ontology/mondo.html     \\
Mouse Pathology Ontology (MPATH)             & https://obofoundry.org/ontology/mpath.html     \\
National Cancer Institute Thesaurus (NCIT)   & https://obofoundry.org/ontology/ncit.html      \\
NCBI Taxonomy Ontology (NCBITaxon)           & https://obofoundry.org/ontology/ncbitaxon.html \\
Neuro Behaviour Ontology (NBO)               & https://obofoundry.org/ontology/nbo.html       \\
Ontology for Biomedical Investigations (OBI) & https://obofoundry.org/ontology/obi.html       \\
Ontology of Medically Related Social Entities (OMRSE)                & https://obofoundry.org/ontology/omrse.html  \\
Population and Community Ontology (PCO)      & https://obofoundry.org/ontology/pco.html       \\
UBERON Multi-species Anatomy Ontology (UBERON)                       & https://obofoundry.org/ontology/uberon.html \\
Unit Ontology (UO)                           & https://obofoundry.org/ontology/uo.html        \\
Vaccine Ontology (VO)                        & https://obofoundry.org/ontology/vo.html        \\ \bottomrule
\end{tabular}%
}
\small
\item $^1$ Vocabulary for fields and terms in the specification have been sourced or mapped to OBO Foundry domain and application ontologies, which are highlighted in this list. New fields and terms for which there were no existing equivalents have been developed and submitted to these ontologies, expanding these community resources
\end{table}