### 2022-05-23 tables/chapter_7/table4.tex

\begin{table}[]
\centering
\caption{A selection of accession numbers of harmonized contextual data records submitted to different public repositories.}
\label{tab:ch7_table4}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
Data   Contributor &
  Repository   Name &
  Accession   Number \\ \midrule
\begin{tabular}[c]{@{}l@{}}African Centre of Excellence for Genomics \\ of Infectious Diseases (Nigeria)\end{tabular} &
  GISAID &
  \begin{tabular}[c]{@{}l@{}}EPI\_ISL\_1035827\\ EPI\_ISL\_1035826\\ EPI\_ISL\_1035825\end{tabular} \\ \midrule
\begin{tabular}[c]{@{}l@{}}COVID-19 Genomic Surveillance \\ Regional Network (Latin America)\end{tabular} &
  GISAID &
  \begin{tabular}[c]{@{}l@{}}EPI\_ISL\_2158821\\ EPI\_ISL\_2158802\\ EPI\_ISL\_2158810\end{tabular} \\ \midrule
\begin{tabular}[c]{@{}l@{}}COVID-19 Genomic Surveillance \\ Regional Network (Latin America)\end{tabular} &
  EMBL-EBI &
  SAMEA8968916 \\ \midrule
\begin{tabular}[c]{@{}l@{}}Rhode Island Department of \\ Health/Broad Institute (SPHERES)\end{tabular} &
  NCBI &
  SAMN18306978 \\ \midrule
\begin{tabular}[c]{@{}l@{}}Massachusetts General Hospital/Broad\\ Institute (SPHERES)\end{tabular} &
  NCBI &
  SAMN18309294 \\ \midrule
Flow Health/Broad Institute (SPHERES) &
  NCBI &
  SAMN18308763 \\ \midrule
\begin{tabular}[c]{@{}l@{}}New Brunswick Diagnostic Virology Reference\\ Center/Public Health Agency of Canada (CanCOGeN)\end{tabular} &
  NCBI &
  SAMN16784832 \\ \midrule
\begin{tabular}[c]{@{}l@{}}Toronto Invasive Bacterial Diseases \\ Network/McMaster University (CanCOGeN)\end{tabular} &
  NCBI &
  SAMN17505317 \\ \midrule
\begin{tabular}[c]{@{}l@{}}Bat coronavirus phylogeography- Université de La Réunion, \\ UMR Processus Infectieux en Milieu Insulaire Tropical (PIMIT) \\ and Field Museum of Natural History\end{tabular} &
  NCBI &
  \begin{tabular}[c]{@{}l@{}}SAMN20400589\\ SAMN20400588\end{tabular}
\end{tabular}%
}
\end{table}