### 2022-05-24 tables/chapter_8/suptable1.tex

\begin{table}[]
\centering
\caption{Software tested during the ASM NGS 2020 hackathon}
\label{tab:ch8_table_sup1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Software Name} &
  \textbf{Software (URL)} &
  \textbf{Test File (URL)} &
  \textbf{Literature Citation (DOI)} \\ \midrule
BUSCO &
  \begin{tabular}[c]{@{}l@{}}https://gitlab.com/\\ ezlab/busco\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/busco.yml\end{tabular} &
  10.1093/bioinformatics/btv351 \\
Centrifuge &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ DaehwanKimLab/centrifuge\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/centrifuge.yml\end{tabular} &
  10.1101/gr.210641.116 \\
CheckM &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ Ecogenomics/CheckM\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/checkm.yml\end{tabular} &
  10.1101/gr.186072.114 \\
chewBBACA &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ B-UMMI/chewBBACA\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/B-UMMI/\\ chewBBACA/blob/master/\\ .github/workflows/chewbbaca.yml\end{tabular} &
  10.1099/mgen.0.000166 \\
CSIS &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ microbinfie-hackathon2020/CSIS\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/CSIS.yml\end{tabular} &
  this manuscript \\
Genotyphi &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ katholt/genotyphi\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/genotyphi.yml\end{tabular} &
  10.1038/ncomms12827 \\
Kraken &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ DerrickWood/kraken\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/kraken.yml\end{tabular} &
  10.1186/gb-2014-15-3-r46 \\
Kraken2 &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ DerrickWood/kraken2\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/kraken2.yml\end{tabular} &
  10.1186/s13059-019-1891-0 \\
KrakenUniq &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ fbreitwieser/krakenuniq\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/krakenuniq.yml\end{tabular} &
  10.1186/s13059-018-1568-0 \\
Pangolin &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ cov-lineages/pangolin\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/pangolin.yml\end{tabular} &
  absent \\
Prokka &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ tseeman/prokka\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/prokka.yml\end{tabular} &
  10.1093/bioinformatics/btu153 \\
Quast &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ ablab/quast\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/quast.yml\end{tabular} &
  10.1093/bioinformatics/btt086 \\
Shovill &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ tseemann/shovill\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/shovill.yml\end{tabular} &
  absent \\
SKESA &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ ncbi/SKESA\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/skesa.yml\end{tabular} &
  10.1186/s13059-018-1540-z \\
Trycycler &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ rrwick/Trycycler\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/trycycler.yml\end{tabular} &
  10.5281/zenodo.4430941 \\
Unicycler &
  \begin{tabular}[c]{@{}l@{}}https://github.com/\\ rrwick/Unicycler\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}https://github.com/microbinfie-\\ hackathon2020/CSIS/blob/main/\\ .github/workflows/unicycler.yml\end{tabular} &
  10.1371/journal.pcbi.1005595 \\ \bottomrule
\end{tabular}%
}
\end{table}