### 2022-05-26 tables/chapter_5/table1.tex

\begin{scriptsize}
\begin{center}

\begin{table}[]
\centering
\caption{Prokaryotic de novo assemblers integrated into LMAS.}
\label{tab:ch5_table1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
Assembler         & Type        & Algorithm                        \\ \midrule
GATBMiniaPipeline & Metagenomic & Multiple   k-mer De Bruijn graph \\
IDBA-UD           & Metagenomic & Multiple   k-mer De Bruijn graph \\
MEGAHIT           & Metagenomic & Multiple   k-mer De Bruijn graph \\
MetaHipMer2       & Metagenomic & Multiple   k-mer De Bruijn graph \\
metaSPAdes        & Metagenomic & Multiple   k-mer De Bruijn graph \\
ABySS             & Genomic     & Single   k-mer De Bruijn graph   \\
BCALM2            & Genomic     & Single   k-mer De Bruijn graph   \\
MINIA             & Genomic     & Single   k-mer De Bruijn graph   \\
SKESA             & Genomic     & Multiple   k-mer De Bruijn graph \\
SPAdes            & Genomic     & Multiple   k-mer De Bruijn graph \\
Unicycler         & Genomic     & Multiple   k-mer De Bruijn graph \\
VelvetOptimizer   & Genomic     & Multiple   k-mer De Bruijn graph \\ \bottomrule
\end{tabular}%
}
\end{table}

\end{center}
\end{scriptsize}