### 2022-06-08 tables/chapter_5/suptable1.tex

\begin{table}[]
\centering
\caption{Tools available for the de novo assembly of prokaryotic genomes. For each tool, its publication is indicated, if available, as well as the assembly algorithm implemented if it was developed explicitly to handle metagenomic datasets. The tools are ordered by the date of the last update, with the source code indicated when available. The tools incorporated in LMAS are indicated as such. }
\label{tab:ch5_suptable1}
Table available for download at \url{https://zenodo.org/record/6623458/files/LMAS\%20Supplemental\%20Material\%20-\%20Tables_Table\%20S1.xlsx}
\end{table}