### 2022-06-08 tables/chapter_5/suptable11.tex

\begin{table}[]
\centering
\caption{Inconsistent contigs produced by the assemblers in 3 LMAS runs. For each assembler, the total number of contigs produced over the 3 runs of the LMAS workflow is indicated, as well as the contigs present in only two and a single run. }
\label{tab:ch5_suptable11}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllll@{}}
\toprule
Assembler &
  Total Contigs &
  \begin{tabular}[c]{@{}l@{}}Contigs present in \\ 3 out of 3 runs\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Contigs present in \\ 2 out of 3 runs\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Contigs present in \\ 1 out of 3 runs\end{tabular} \\ \midrule
ABySS             & 7638  & 2505 & 50  & 23  \\
BCALM2            & 1962  & 654  & 0   & 0   \\
GATBMiniaPipeline & 14321 & 4753 & 25  & 12  \\
IDBA-UD           & 15383 & 5115 & 16  & 6   \\
MEGAHIT           & 5440  & 1812 & 1   & 2   \\
MetaHipMer2       & 2358  & 772  & 16  & 10  \\
metaSPAdes        & 6462  & 2154 & 0   & 0   \\
minia             & 6427  & 2106 & 41  & 27  \\
SKESA             & 11187 & 3729 & 0   & 0   \\
SPAdes            & 8631  & 2877 & 0   & 0   \\
Unicycler         & 5598  & 1866 & 0   & 0   \\
VelvetOptimiser   & 2315  & 553  & 112 & 432 \\
\bottomrule
\end{tabular}%
}
\end{table}