### 2022-06-08 tables/chapter_5/suptable12.tex

\begin{table}[]
\centering
\caption{Global assembly metrics for single and multiple k-mer dBg assemblers. The median and the minimum and maximum values obtained are presented for each metric for all samples in 3 runs of LMAS. Single k-mer bBg assemblers: ABySS, BCALM2 and minia. Multiple k-mer bBg assembler: GATBMiniaPipeline, IDBA-UD, MEGAHIT, MetaHipMer2, metaSPAdes, SKESA, SPAdes, Unicycler and VelverOptimiser.}
\label{tab:ch5_suptable12}
Table available for download at \url{https://zenodo.org/record/6623458/files/LMAS\%20Supplemental\%20Material\%20-\%20Tables_Table\%20S12.xlsx}
\end{table}