### 2022-06-08 tables/chapter_5/suptable13.tex

\begin{table}[]
\centering
\caption{Global assembly metrics for genomic and metagenomic multiple k-mer dBg assemblers. The median and the minimum and maximum values obtained are presented for each metric for all samples in 3 runs of LMAS. Genomic assemblers: SKESA, SPAdes and Unicycler. Metagenomic assemblers: GATBMiniaPipeline, IDBA-UD, MEGAHIT and metaSPAdes.}
\label{tab:ch5_suptable13}
Table available for download at \url{https://zenodo.org/record/6623458/files/LMAS\%20Supplemental\%20Material\%20-\%20Tables_Table\%20S13.xlsx}
\end{table}