### 2022-06-08 tables/chapter_5/suptable3.tex

\begin{table}[]
\centering
\caption{The ZymoBIOMICS Microbial Community Standard datasets. Set of raw sequence reads used as input in LMAS  of mock communities with an even and logarithmic distribution of species, from real sequencing runs and simulated read datasets, with and without error, matching the intended distribution of species in ZymoBIOMICS Microbial Community Standard.}
\label{tab:ch5_suptable3}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lllll@{}}
\toprule
Sample   Name & Distribution & Error Model         & Read Length (bp) & Read Pairs (M) \\ \midrule
ENN           & Even         & None                & 150              & 8,5            \\
EMS           & Even         & Illumina   MiSeq    & 150              & 8,5            \\
ERR2984773    & Even         & Real   MiSeq Sample & 150              & 8,5            \\
LNN           & Log          & None                & 100              & 47,5           \\
LHS           & Log          & Illumina   HiSeq    & 100              & 47,5           \\
ERR2935805    & Log          & Real   HiSeq Sample & 100              & 47,5           \\ \bottomrule
\end{tabular}%
}
\end{table}