### 2022-06-08 tables/chapter_5/suptable15.tex
Updates from Overleaf

\begin{table}[]
\centering
\caption{Per reference quality metrics variation in three LMAS s for sample EMS per assembler of the ZymoBIOMICS microbial community standard dataset. The average calculated for all samples in the dataset for the 3 independent LMAS runs, followed by the minimum and maximum values obtained, are presented for each metric for each assembler.}
\label{tab:ch5_suptable15}
Table available for download at \url{https://zenodo.org/record/6623458/files/LMAS\%20Supplemental\%20Material\%20-\%20Tables_Table\%20S15.xlsx}
\end{table}