### 2022-06-08 tables/chapter_5/suptable17.tex

\begin{scriptsize}
\begin{center}

\begin{table}[]
\centering
\caption{Inconsistent gaps produced by the assemblers in 3 LMAS runs. For each assembler, the total number of gaps consistently produced in relation to the reference replicons over the 3 runs of the LMAS workflow is indicated, as well as gaps present in only two and a single run.}
\label{tab:ch5_suptable17}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
Assembler &
  \begin{tabular}[c]{@{}l@{}}Gaps present in \\ 3 out of 3 runs\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Gaps present in \\ 2 out of 3 runs\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Gaps present in \\ 1 out of 3 runs\end{tabular} \\ \midrule
GATBMiniaPipeline & 3108 & 278 & 157 \\
IDBA-UD           & 4458 & 68  & 40  \\
MEGAHIT           & 1017 & 122 & 89  \\
metaSPAdes        & 1185 & 0   & 0   \\
SKESA             & 6681 & 0   & 0   \\
SPAdes            & 2433 & 0   & 0   \\
Unicycler         & 1512 & 0   & 0  \\\bottomrule
\end{tabular}%
}
\end{table}

\end{center}
\end{scriptsize}