### 2022-06-08 tables/chapter_5/suptable19.tex

\begin{scriptsize}
\begin{center}

\begin{table}[]
\centering
\caption{Number of tRNA and rRNA coding sequencing, and mobile elements in ZymoBIOMICS microbial community standard reference replicons. The average calculated for all samples in the dataset for the 3 independent LMAS runs, followed by the minimum and maximum values obtained, are presented for each metric for each assembler. }
\label{tab:ch5_suptable19}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
Reference               & tRNA & rRNA & Mobile elements \\\midrule
\textit{Bacillus subtilis}       & 86   & 31   & 17              \\
\textit{Enterococcus faecalis}   & 62   & 12   & 7               \\
\textit{Escherichia coli}        & 89   & 22   & 85              \\
\textit{Lactobacillus fermentum} & 59   & 15   & 80              \\
\textit{Listeria monocytogenes}  & 67   & 18   & 3               \\
\textit{Pseudomonas aeruginosa}  & 78   & 12   & 19              \\
\textit{Salmonella enterica}     & 88   & 22   & 39              \\
\textit{Staphylococcus aureus}   & 60   & 19   & 7              \\\bottomrule
\end{tabular}%
}
\end{table}

\end{center}
\end{scriptsize}