### 2022-06-08 tables/chapter_5/suptable20.tex

\begin{table}[]
\centering
\caption{Taxonomic classification of the ZymoBIOMICS microbial community standard dataset. The classification was performed with Kraken2, using the Standard Database. The results are presented as the percentage of classified reads for the 8 bacterial species in the community, as well as unclassified reads and the group of reads that are classified as species not contained in the community standard.}
\label{tab:ch5_suptable20}
Table available for download at \url{https://zenodo.org/record/6623458/files/LMAS\%20Supplemental\%20Material\%20-\%20Tables_Table\%20S20.xlsx}
\end{table}