### 2022-11-03 chapters/acknowledgements.tex
add hAMRonization paper

There are several key features necessary for the successful completion of a doctoral dissertation work: perseverance, critical spirit, passion, and a healthy dose of stubbornness. But none of these matter if a support network is not available. Throughout the five years that this project has dominated my life, both in and outside of the lab, several key-people have been crucial so that I could be here today, writing the acknowledgement section of my life's work at the end of a very long road.

To my supervisors, Dr. João Carriço and Prof. Dr. Mário Ramirez, my deepest thanks for being not only a continuous source of inspiration, but of unconditional support. At the very beginning of this journey, when this was not more than just an idea, you've embraced me wanting to "change paths" and switch from comparative genomics to the wonderful and exciting world of metagenomics. Some of the initial excitement has faded, as it became more evident that this was a challenging road, but with a healthy dose of scepticism you've helped me keep my feet on the ground while my head stayed in the clouds, forever passionate about embracing new challenges and taking on new adventures. In particular to Mário, who started out as my mentor and had to step up to the challenge of supervising me, I give my deepest thanks. Without your support I would not be the scientist I am today, and I do hope one day to make you proud. I will forever carry with me the example of what is to lead a team with grace and patience. I would also like to thank Prof. Dr. John Rossen, who albeit briefly, was part of my original advisory team and thanks to whom I was able to give my first steps in the world of metagenomics.

To the MRamirez lab colleagues, both current and past, you all have marked me and my journey in some way. For all the moments shared, all the laughter and the many shoulders to cry on, I thank you. To Miguel Machado and Catarina Pato, you guys have been my rock. To all the "binfies" who crossed my path, Bruno, Mickael, Diogo, Tiago, and my dear "pitufos" Pedro and Rafael, I've learned so much with each and every one of you. It has been a pleasure and a privilege. Last but not least, to Joana, who showed me the value of compassion, of dedication and of true friendship.