### 2022-11-04 chapters/acknowledgements.tex
finish first draft of the thank you section

There are several key features necessary for the successful completion of a doctoral dissertation work: perseverance, critical spirit, passion, and a healthy dose of stubbornness. But none of these matter if a support network is not available. Throughout the five years that this project has dominated my life, both in and outside of the lab, several key-people have been crucial so that I could be here today, writing the acknowledgement section of my life's work at the end of a very long road.

To my supervisors, Dr. João Carriço and Prof. Dr. Mário Ramirez, my deepest thanks for being not only a continuous source of inspiration, but of unconditional support. At the very beginning of this journey, when this was not more than just an idea, you've embraced me wanting to "change paths" and switch from comparative genomics to the wonderful and exciting world of metagenomics. Some of the initial excitement has faded, as it became more evident that this was a challenging road, but with a healthy dose of scepticism you've helped me keep my feet on the ground while my head stayed in the clouds, forever passionate about embracing new challenges and taking on new adventures. In particular to Mário, who started out as my mentor and had to step up to the challenge of supervising me, I give my deepest thanks. Without your support I would not be the scientist I am today, and I do hope one day to make you proud. I will forever carry with me the example of what is to lead a team with grace and patience. I would also like to thank Prof. Dr. John Rossen, who albeit briefly, was part of my original advisory team and thanks to whom I was able to give my first steps in the world of metagenomics.

To the MRamirez lab colleagues, both current and past, you all have marked me and my journey in some way. For all the moments shared, all the laughter and the many shoulders to cry on, I thank you. To Miguel Machado and Catarina Pato, you guys have been my rock. To all the "binfies" who crossed my path, Bruno, Mickael, Diogo, Tiago, and my dear "pitufos" Pedro and Rafael, I've learned so much with each and every one of you. It has been a pleasure and a privilege. Last but not least, to Joana, who showed me the value of compassion, of dedication and of true friendship.

To the friends I made in the Netherlands, Casper, Lieke, Boas, and whom I had the pleasure of sharing a fantastic and dynamic workspace with, Erley, Nilay, Peppe, Leo and Natacha, my biggest thanks. You made my adventure of living abroad memorable.

To my friends who have become more than family, Nuno, Alicia, Sofia, Neto, Margarida, thank you for being there in my lowest and highest moments. I'm blessed to have you in my life. To you, Rúben, who grew from a little pebble in my shoe to a huge rock for me to lie against, thank you.

To my mum and dad, thank you for your unconditional love, for being my biggest fans, my most loyal allies. To my sister, who never fails to push me to be better, you're my best friend in the whole world. To Miguel e Inês, you bring so much joy to my life. Being your aunt is not only a pleasure, it's an honour to be able to watch you grow. Even though work has made me miss many birthdays over the years... To my brother in law Nuno, thank you for showing me what perseverance means. To my grandma, who is more than a mother to me, and my grandpa, who might not be here anymore but will live forever in my heart, thank you for showing me what love means, now and forever. To you, my dear grandpa, I dedicate my life's work, to honour you and your memory every day of my life. O primeiro bebe-se inteiro.