### 2023-01-23 main.tex
Add sections to thesis body

%-----------------------------------------------------------------
% Preamble
%------------------------------------------------------------------
\documentclass[12pt,a4paper,twoside,openright]{book}
\usepackage{mathptmx} % times no roman-like font
\usepackage[left=30mm, right=25mm, top=25mm, bottom=25mm]{geometry} % margins
\usepackage{setspace}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\baselinestretch}{1.15}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{figures/}}
\def\blankpage{%
      \clearpage%
      \thispagestyle{empty}%
      \null%
      \clearpage}

%-----------------------------------------------------------------
% Packages
%------------------------------------------------------------------
\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\thechapter. #1 }}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\usepackage{titlesec}
\fancyhf{}
\fancyhead[RO]{\bfseries\rightmark}
\fancyhead[LE]{\bfseries\leftmark}
\setlength{\headheight}{15pt}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\headheight}{0.5pt}
\fancypagestyle{plain}{
  \fancyhead{}
  \renewcommand{\headrulewidth}{0pt}
}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[nottoc]{tocbibind}
\usepackage[backend=biber,style=numeric-comp,maxcitenames=1,autocite=inline,natbib=true,sorting=none]{biblatex}
\addbibresource{references.bib}
\usepackage{indentfirst}
\setlength{\parskip}{1em}
\setcounter{tocdepth}{4}
\setcounter{secnumdepth}{4}
\usepackage{enumitem}
\usepackage[font=footnotesize]{caption}
\usepackage[toc,page]{appendix}
\usepackage{pdfpages}
\usepackage{fmtcount}
\usepackage{url}
\usepackage{amsmath}
\usepackage{bm}
\usepackage[utf8]{inputenc}
\usepackage{filecontents}
\usepackage{float}
\raggedbottom
\usepackage{subcaption}
\usepackage{eurosym}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage[acronym]
\makeatletter

\def\cleardoublepage{\clearpage\if@twoside \ifodd\c@page\else
    \hbox{}
    \thispagestyle{plain}
    \newpage
    \if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother \clearpage{\pagestyle{plain}\cleardoublepage}

\providecommand{\keywords}[1]{\textbf{Keywords:} #1}
\providecommand{\keywordsP}[1]{\textbf{Palavras chave:} #1}
\newcommand{\secref}[1]{\autoref{#1}. \nameref{#1}}
\usepackage{lscape}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{multirow}
\hypersetup{hidelinks}
\usepackage{longtable}
\usepackage{acronym} 
\usepackage{amssymb}
\usepackage{pifont}
\usepackage[symbol]{footmisc}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%-----------------------------------------------------------------
% Begin Document
%------------------------------------------------------------------

\begin{document}
\begin{sloppy}
\frontmatter

%-----------------------------------------------------------------
% Cover
%------------------------------------------------------------------
\input{titlepage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Folha de rosto
%------------------------------------------------------------------
\input{coverpage}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Disclaimer
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{\fill}
As opiniões expressas nesta publicação são da exclusiva responsabilidade do seu autor.
\blankpage

%-----------------------------------------------------------------
% Dedication
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\vspace*{3cm}
\begin{center}
\emph{"The greatest adventure is what lies ahead.\\
Today and tomorrow are yet to be said.\\
The chances, the changes are all yours to make.\\
The mould of your life is in your hands to break."} \\
\end{center} 
\begin{flushright}-J. R. R. Tolkien, The Hobbit\end{flushright}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Acknowledgements
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Acknowledgements}
\chapter*{Acknowledgements}
\input{chapters/acknowledgements.tex}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Summary
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Summary}
\chapter*{Summary}
\input{chapters/abstract.tex}

\keywords{Clinical Microbiology, Genome Sequencing, Metagenomics, Reproducibility, Interoperability}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Resumo
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Resumo}
\chapter*{Resumo}
\input{chapters/resumo.tex}

\keywords{Microbiologia Clínica, Sequênciação genómica, Metagenómica, Reproducibilidade, Interoperabilidade}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Thesis Outline
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Thesis Outline}
\chapter*{Thesis Outline}
\input{chapters/thesis_outline}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Abbreviations
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Abbreviation}
\chapter*{Abbreviation}
\input{chapters/acconyms}
\clearpage \thispagestyle{empty}\mbox{}\clearpage

%-----------------------------------------------------------------
% Table of Contents
%------------------------------------------------------------------
\newpage
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{Table of Contents}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoftables
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\thispagestyle{plain}
\listoffigures
%\clearpage \thispagestyle{empty}\mbox{}\clearpage

\mainmatter
%\pagenumbering{arabic} %numeros em numeração árabe

%-----------------------------------------------------------------
% Introduction
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{General Introduction\label{ch:introduction}}
\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/introduction}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

\part{Applying metagenomics in the clinical context}
%-----------------------------------------------------------------
% Paper 1 - Natacha
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{Critical steps in clinical shotgun metagenomics for the concomitant detection and typing of microbial pathogens\label{ch:paper1}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_2}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

%-----------------------------------------------------------------
% Paper 2 - Peppe
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{Detection of a novel \textit{mcr-5.4} gene variant in hospital tap water by shotgun metagenomic sequencing\label{ch:paper2}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_3}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

%-----------------------------------------------------------------
% Paper 3 - DEN-IM
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{DEN-IM: Dengue virus genotyping from shotgun and targeted metagenomics\label{ch:paper3}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_4}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

\part{Impact of \textit{de novo} assemblers in metagenomics}
%-----------------------------------------------------------------
% Paper 4 - LMAS
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{LMAS: Last Metagenomic Assembler Standing\label{ch:paper4}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_5}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

\part{Challenges of data availability in metagenomics and beyond}
%-----------------------------------------------------------------
% Paper 5 - hAMRonization
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{hAMRonization: Enhancing antimicrobial resistance prediction using PHA4GE standards and specification\label{ch:paper5}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_6}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

%-----------------------------------------------------------------
% Paper 6 - SARS-CoV-2
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{Future-proofing and maximising the utility of metadata: The PHA4GE SARS-CoV-2 contextual data specification package\label{ch:paper6}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_7}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

\part{Crowdsourcing for software robustness in metagenomics and beyond}
%-----------------------------------------------------------------
% Paper 7 - Software Testing
%------------------------------------------------------------------
\newpage
\thispagestyle{empty}
\chapter{Software testing in microbial bioinformatics: a call to action\label{ch:paper7}}

\thispagestyle{empty}
\clearpage \thispagestyle{empty}\mbox{}\clearpage
\newpage
\begin{refsection}
\include{chapters/chapter_8}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

%-----------------------------------------------------------------
% Discussion
%-----------------------------------------------------------------

\newpage
\thispagestyle{empty}
\chapter{General Discussion \label{ch:discussion}}

\clearpage \thispagestyle{empty}\mbox{}\clearpage
\begin{refsection}
\include{chapters/discussion}
\newpage
\section{References}
\printbibliography[heading=none]
\end{refsection}

%-----------------------------------------------------------------
% Conclusion
%-----------------------------------------------------------------
\newpage
\chapter{Conclusion \label{ch:conclusion}}

\clearpage \thispagestyle{empty}\mbox{}\clearpage
\begin{conclusion}
\include{chapters/conclusion}
\newpage
\section{References}
\printbibliography[heading=none]
\end{conclusion}

%-----------------------------------------------------------------
% Appendix
%-----------------------------------------------------------------
%\appendix
\newpage
\chapter{Appendix}
\includepdf[pages={1-}]{resources/papers/s41598-018-31873-w.pdf}
\includepdf[pages={1-}]{resources/papers/dkz363.pdf}
\includepdf[pages={1-}]{resources/papers/mgen000328.pdf}
\includepdf[pages={1-}]{resources/papers/giac122.pdf}
\includepdf[pages={1-}]{resources/papers/giac003.pdf}
\includepdf[pages={1-}]{resources/papers/mgen000790.pdf}

\end{sloppy}
\end{document}